---
title: "assign4"
author: "Christopher Huong"
date: "`r Sys.Date()`"
output: html_document
---


```{r message = F}
library(tidyverse)
library(ggplot2)

# library(haven)
# dat <- read_xpt('LLCP2021.XPT_')
# 
# save(dat, file = "BRFSS.RData")

load('BRFSS.RData')



d <- dat %>%
  select("ADDEPEV3", "_AGEG5YR", "_SEX", "_RACE")


colnames(d) <- c("DEPR", "AGE", "SEX", "RACE")

```
Choose variables, rename columns

```{r}
d <- d %>%
  mutate(DEPR = case_when(DEPR == 1 ~ 1, # 1 = yes depr
                          DEPR == 2 ~ 0, # 0 = no depr
                          DEPR == 7 ~ NA_real_,
                          DEPR == 9 ~ NA_real_))



d[, 2][d[, 2] == 14] <- NA #assign 14 to NA in age column


d$SEX <- factor(d$SEX, labels = c("M", "F"))


d <- d %>%
  mutate(RACE = case_when(RACE == 1 ~ "White", 
                          RACE == 2 ~ "Black",
                          RACE == 3 ~ "Native",
                          RACE == 4 ~ "Asian",
                          RACE == 5 ~ "PI",
                          RACE == 6 ~ NA_character_,
                          RACE == 7 ~ NA_character_,
                          RACE == 8 ~ "Hispanic",
                          RACE == 9 ~ NA_character_
  ))

d$RACE <- as.factor(d$RACE)
d$RACE <- relevel(d$RACE, ref = "White")

```
Cleaning

```{r}
mod <- glm(DEPR ~ RACE + SEX + AGE,
            family = "binomial",
            data = d)

summary(mod)
exp(cbind(OR = coef(mod), confint(mod)))
```
Modeling. All terms are significantly associated with probability of reporting having been diagnosed with depression







# part2 -------------------------------------------------------------------



```{r}
df <- dat %>%
  select("ADDEPEV3", "_AGEG5YR", "_SEX", "_STATE",
  "ACEDEPRS", "ACEDRINK", "ACEDRUGS", "ACEPRISN",
  "ACEDIVRC", "ACEPUNCH", "ACEHURT1", "ACESWEAR", "ACETOUCH",
  "ACETTHEM", "ACEHVSEX",  "_TOTINDA", "MENTHLTH", "_EDUCAG", 
  "_RACE", "_BMI5CAT", "_INCOMG1"
  
    
  )


colnames(df) <- c("DEPR", "AGE", "SEX", "STATE",
  "ACEDEPRS", "ACEDRINK", "ACEDRUGS", "ACEPRISN",
  "ACEDIVRC", "ACEPUNCH", "ACEHURT1", "ACESWEAR", "ACETOUCH",
  "ACETTHEM", "ACEHVSEX", "PA", "MH_DAYS", "EDU",
  "RACE", "BMI", "INCOME"
  
  )
```
Import and rename


```{r}

df <- df %>%
  mutate(DEPR = case_when(DEPR == 1 ~ 1, # 1 = yes depr
                          DEPR == 2 ~ 0, # 0 = no depr
                          DEPR == 7 ~ NA_real_,
                          DEPR == 9 ~ NA_real_))


df <- df %>%
  mutate(PA = case_when(PA == 1 ~ 1, # 1 = yes PA
                        PA == 2 ~ 0, # 0 = no PA
                        PA == 9 ~ NA_real_))




df <- df %>%
  mutate(EDU = case_when(EDU == 1 ~ "NO_HS",
                         EDU == 2 ~ "HS",
                         EDU == 3 ~ "ATTENDCOL",
                         EDU == 4 ~ "GRADCOL",
                         EDU == 9 ~ NA_character_))


df <- df %>%
  mutate(RACE = case_when(RACE == 1 ~ "WHITE",
                          RACE == 2 ~ "BLACK",
                          RACE == 3 ~ "NATIVE",
                          RACE == 4 ~ "ASIAN",
                          RACE == 5 ~ "ISLANDER",
                          RACE == 6 ~ "OTHER",
                          RACE == 7 ~ "MULTI",
                          RACE == 8 ~ "HISPANIC",
                          RACE == 9 ~ NA_character_))

df <- df %>%
  mutate(BMI = case_when(BMI == 1 ~ "UNDERW",
                         BMI == 2 ~ "NORMW",
                         BMI == 3 ~ "OVERW",
                         BMI == 4 ~ "OBESE"))


df <- df %>%
  mutate(INCOME = case_when(INCOME == 1 ~ "0-15K",
                            INCOME == 2 ~ "15-25K",
                            INCOME == 3 ~ "25-35K",
                            INCOME == 4 ~ "35-50K",
                            INCOME == 5 ~ "50-100K",
                            INCOME == 6 ~ "100-200K",
                            INCOME == 7 ~ "200K+",
                            INCOME == 9 ~ NA_character_))


df[, 2][df[, 2] == 14] <- NA #assign 14 to NA in age column
```
Recode

```{r}
# convert to factors and assign reference levels
df$SEX <- factor(df$SEX, labels = c("M", "F"))
df$EDU <- as.factor(df$EDU)
df$EDU <- relevel(df$EDU, ref = "NO_HS")
df$RACE <-as.factor(df$RACE)
df$RACE <- relevel(df$RACE, ref = "WHITE")
df$BMI <- as.factor(df$BMI)
df$BMI <- relevel(df$BMI, ref = "NORMW")
df$INCOME <- factor(df$INCOME)
df$INCOME <- relevel(df$INCOME, ref = "0-15K")


# replace other answers with NA
df[, 5:15][df[, 5:15] == 7] <- NA
df[, 5:15][df[, 5:15] == 8] <- NA
df[, 5:15][df[, 5:15] == 9] <- NA


# ACEDEPRS - ACEDIVRC: yes=1, no=2
#convert to yes=1, no=0
df[, 5:9][df[, 5:9] == 2] <- 0 
# ACEHURT1 - ACEHVSEX: never=1, once=2, morethanonce=3
#convert to never=0, once=1, morethanonce=1
df[, 10:15][df[, 10:15] == 1] <- 0
df[, 10:15][df[, 10:15] == 2] <- 1
df[, 10:15][df[, 10:15] == 3] <- 1


# 7 = don't know/not sure, 9 = refused. convert all to missing


df <- na.omit(df)
sum(is.na(df))
#remove all NA, 40024 rows left, ~9% of original sample
```
More recoding



```{r}
df <- df %>%
  mutate(ACE = ACEDEPRS + ACEDRINK + ACEDRUGS + ACEPRISN +
              ACEDIVRC + ACEPUNCH + ACEHURT1 + ACESWEAR + 
              ACETOUCH + ACETTHEM + ACEHVSEX)

# create new column for total ACEs (0-11)



df <- df %>%
  mutate(ACE_PSYCH = if_else(ACESWEAR == 1, #psychological abuse
                            1L, 0L))

df <- df %>%
  mutate(ACE_PHYS = if_else(ACEHURT1 == 1, #physical abuse
                            1L, 0L))

df <- df %>%
  mutate(ACE_SXL = if_else(ACETOUCH == 1 | #sexual abuse
                           ACETTHEM == 1 |
                           ACEHVSEX == 1,
                           1L, 0L))
            
df <- df %>%
  mutate(ACE_HD_SA = if_else(ACEDRINK == 1 | #household dysf, substance abuse
                             ACEDRUGS == 1,
                             1L, 0L))
              
df <- df %>%
  mutate(ACE_HD_MI = if_else(ACEDEPRS == 1, #household dysf, mental illness
                             1L, 0L))

df <- df %>%
  mutate(ACE_HD_PAR = if_else(ACEDIVRC == 1 | #household dysf, parental dispute
                              ACEPUNCH == 1,
                              1L, 0L))

df <- df %>%
  mutate(ACE_HD_CRIM = if_else(ACEPRISN == 1, #household dysf, prison
                               1L, 0L))

df <- df %>%
  mutate(ACE_CAT_SUM = ACE_PSYCH + ACE_PHYS + ACE_SXL + ACE_HD_SA
                     + ACE_HD_MI + ACE_HD_PAR + ACE_HD_CRIM)
# create new column total ACE category score (0-7) per original ACE study
```

```{r}
library(fishmethods)

clus.rho(popchar = df$DEPR, cluster = df$STATE)
clus.rho(popchar = df$PA, cluster = df$STATE)
```
Calculate ICC to determine if need to cluster by state
Within state variance far outweights between state variance, no need to cluster




Predicting PA
```{r}
mod0 <- glm(as.factor(PA) ~ ACE_CAT_SUM  + DEPR 
            + AGE + SEX + EDU + RACE + BMI + INCOME, 
            family = "binomial", data = df)

summary(mod0)
exp(cbind(OR = coef(mod0), confint(mod0)))
```
ACE score does not predict PA


Predicting diagnosis of depression with ACE score and PA
```{r}
mod1 <- glm(DEPR ~ ACE_CAT_SUM + as.factor(PA), 
            family = "binomial", data = df)

summary(mod1)
exp(cbind(OR = coef(mod1), confint(mod1)))
```
Compute adjusted ORs
Increase in ACE category = 1.47x likely to report depression diagnosis
Reporting PA = 59% as likely to report depression diagnosis

```{r}
mod2 <- glm(DEPR ~ ACE_CAT_SUM + as.factor(PA)
            + AGE + SEX + EDU + RACE + BMI + INCOME, 
             family = "binomial", data = df)

summary(mod2)
exp(cbind(OR = coef(mod2), confint(mod2)))[c(2:11,16,19:24),]
```
Compute adjusted ORs, extract estimates with p < 0.01
Estimates are relatively stable when controlling for covariates


Model ACE*PA interaction
```{r}
mod3 <- glm(DEPR ~ ACE_CAT_SUM*as.factor(PA)
          + AGE + SEX + EDU + RACE + BMI + INCOME, 
            family = "binomial", data = df)

summary(mod3)
```
Nonsignificant interaction term

```{r}
BIC(mod1, mod2, mod3)
```
Model 2 with the set of covariates has a lower BIC and therefore is preferable to model 1 and 3



Look at individual ACE categories
```{r}
cols <- c(23:29)
df <- df %>% mutate_at(cols, factor)

mod4 <- glm(DEPR ~ ACE_PSYCH +
              ACE_PHYS +
              ACE_SXL +
              ACE_HD_SA +
              ACE_HD_MI +
              ACE_HD_PAR +
              ACE_HD_CRIM +
              PA + AGE + SEX + EDU + RACE + BMI + INCOME,
            family = "binomial", data = df)

summary(mod4)                  
exp(cbind(OR = coef(mod4), confint(mod4)))[c(2:8),]       
              
```


```{r}
ggplot(df, aes(x=as.factor(ACE_CAT_SUM), fill=as.factor(DEPR)))+
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="dodge" ) +
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
            stat="count", position=position_dodge(0.9), vjust=-0.5)+
  ylab('IDK') +
  scale_y_continuous(labels = scales::percent)


```





