---
title: "assign5"
output: html_document
date: "2023-02-27"
---
```{r, message=F}
library(tidyverse)
library(ggplot2)
library(margins)
library(haven)
library(janitor)

```

```{r}
load('BRFSS.RData')





df <- dat %>%
  select("ADDEPEV3", "_AGEG5YR", "_SEX", "_STATE",
         "ACEDEPRS", "ACEDRINK", "ACEDRUGS", "ACEPRISN",
         "ACEDIVRC", "ACEPUNCH", "ACEHURT1", "ACESWEAR", "ACETOUCH",
         "ACETTHEM", "ACEHVSEX",  "_TOTINDA", "MENTHLTH", "_EDUCAG", 
         "_RACE", "_BMI5CAT", "_INCOMG1"
         
         
  )



colnames(df) <- c("DEPR", "AGE", "SEX", "STATE",
                  "ACEDEPRS", "ACEDRINK", "ACEDRUGS", "ACEPRISN",
                  "ACEDIVRC", "ACEPUNCH", "ACEHURT1", "ACESWEAR", "ACETOUCH",
                  "ACETTHEM", "ACEHVSEX", "PA", "MH_DAYS", "EDU",
                  "RACE", "BMI", "INCOME"
                  
)
```
# DEPR
# Question: (Ever told) (you had) a depressive disorder 
# (including depression, major depression, dysthymia, or minor depression)? 


# PA
# Question: Adults who reported doing physical activity or exercise 
# during the past 30 days other than their regular job 





#Input NA for coded missing
```{r}
df <- df %>%
  mutate(DEPR = case_when(DEPR == 1 ~ 1, # 1 = yes depr
                          DEPR == 2 ~ 0, # 0 = no depr
                          DEPR == 7 ~ NA_real_,
                          DEPR == 9 ~ NA_real_))


df[, 17][df[, 17] == 88] <- 0 #assign 88 to 0 in MH_DAYS column
df[, 17][df[, 17] == 77] <- NA
df[, 17][df[, 17] == 99] <- NA 


df <- df %>%
  mutate(PA = case_when(PA == 1 ~ 1, # 1 = yes PA
                        PA == 2 ~ 0, # 0 = no PA
                        PA == 9 ~ NA_real_))




df <- df %>%
  mutate(EDU = case_when(EDU == 1 ~ "NO_HS",
                         EDU == 2 ~ "HS",
                         EDU == 3 ~ "ATTENDCOL",
                         EDU == 4 ~ "GRADCOL",
                         EDU == 9 ~ NA_character_))


df <- df %>%
  mutate(RACE = case_when(RACE == 1 ~ "WHITE",
                          RACE == 2 ~ "BLACK",
                          RACE == 3 ~ "NATIVE",
                          RACE == 4 ~ "ASIAN",
                          RACE == 5 ~ "ISLANDER",
                          RACE == 6 ~ "OTHER",
                          RACE == 7 ~ "MULTI",
                          RACE == 8 ~ "HISPANIC",
                          RACE == 9 ~ NA_character_))

df <- df %>%
  mutate(BMI = case_when(BMI == 1 ~ "UNDERW",
                         BMI == 2 ~ "NORMW",
                         BMI == 3 ~ "OVERW",
                         BMI == 4 ~ "OBESE"))


df <- df %>%
  mutate(INCOME = case_when(INCOME == 1 ~ "0-15K",
                            INCOME == 2 ~ "15-25K",
                            INCOME == 3 ~ "25-35K",
                            INCOME == 4 ~ "35-50K",
                            INCOME == 5 ~ "50-100K",
                            INCOME == 6 ~ "100-200K",
                            INCOME == 7 ~ "200K+",
                            INCOME == 9 ~ NA_character_))



df[, 2][df[, 2] == 14] <- NA #assign 14 to NA in age column

# convert to factors and assign reference levels
df$SEX <- factor(df$SEX, labels = c("M", "F"))
df$EDU <- as.factor(df$EDU)
df$EDU <- relevel(df$EDU, ref = "NO_HS")
df$RACE <-as.factor(df$RACE)
df$RACE <- relevel(df$RACE, ref = "WHITE")
df$BMI <- as.factor(df$BMI)
df$BMI <- relevel(df$BMI, ref = "NORMW")
df$INCOME <- factor(df$INCOME)
df$INCOME <- relevel(df$INCOME, ref = "0-15K")






# replace other answers with NA in ACE Qs
df[, 5:15][df[, 5:15] == 7] <- NA
df[, 5:15][df[, 5:15] == 8] <- NA
df[, 5:15][df[, 5:15] == 9] <- NA




# ACEDEPRS - ACEDIVRC: yes=1, no=2
#convert to yes=1, no=0
df[, 5:9][df[, 5:9] == 2] <- 0 
# ACEHURT1 - ACEHVSEX: never=1, once=2, morethanonce=3
#convert to never=0, once=1, morethanonce=1
df[, 10:15][df[, 10:15] == 1] <- 0
df[, 10:15][df[, 10:15] == 2] <- 1
df[, 10:15][df[, 10:15] == 3] <- 1


# 7 = don't know/not sure, 9 = refused. convert all to missing


df <- na.omit(df)
sum(is.na(df))
#remove all NA, 39632 rows left, ~9% of original sample


df <- df %>%
  mutate(ACE_PSYCH = if_else(ACESWEAR == 1, #psychological abuse
                             1L, 0L))

df <- df %>%
  mutate(ACE_PHYS = if_else(ACEHURT1 == 1, #physical abuse
                            1L, 0L))

df <- df %>%
  mutate(ACE_SXL = if_else(ACETOUCH == 1 | #sexual abuse
                             ACETTHEM == 1 |
                             ACEHVSEX == 1,
                           1L, 0L))

df <- df %>%
  mutate(ACE_HD_SA = if_else(ACEDRINK == 1 | #household dysf, substance abuse
                               ACEDRUGS == 1,
                             1L, 0L))

df <- df %>%
  mutate(ACE_HD_MI = if_else(ACEDEPRS == 1, #household dysf, mental illness
                             1L, 0L))

df <- df %>%
  mutate(ACE_HD_PAR = if_else(ACEDIVRC == 1 | #household dysf, parental dispute
                                ACEPUNCH == 1,
                              1L, 0L))

df <- df %>%
  mutate(ACE_HD_CRIM = if_else(ACEPRISN == 1, #household dysf, prison
                               1L, 0L))

df <- df %>%
  mutate(ACE_CAT_SUM = ACE_PSYCH + ACE_PHYS + ACE_SXL + ACE_HD_SA
         + ACE_HD_MI + ACE_HD_PAR + ACE_HD_CRIM)
#total ACE category score (0-7)
```





# analysis ----------------------------------------------------------------

```{r}
m1 <- glm(DEPR ~ ACE_CAT_SUM + PA + AGE + SEX + EDU,
          data = df,
          family = binomial(logit))


summary(m1)
exp(cbind(OR = coef(m1), confint(m1)))
```
Basic logistic regression, and exponentiating the coefficients for aORs
```{r}
me1 <- margins(m1, variables = NULL)
summary(me1)
```
Average marginal effects
Each increase in ace score increases probability of depression by ~5%
PA decreases probability of depression by ~8%
Being female increases by 10%

```{r}
margins(m1, at = list(ACE_CAT_SUM = range(df$ACE_CAT_SUM)) )
```

```{r}
m2 <- glm(DEPR~ ACE_CAT_SUM*SEX + PA + AGE + EDU,
          data = df,
          family = binomial(logit))

summary(m2)
exp(cbind(OR = coef(m2), confint(m2)))
```

Model interaction term for ACE*SEX to see if probability of DEPR as a function of ACE score differs by SEX
```{r}
me2 <- margins(m2, variables = NULL)
summary(me2)
```
Significant interaction term.
AME's seem to be unchanged when adding interaction

```{r}
margins(m2, at = list(ACE_CAT_SUM = range(df$ACE_CAT_SUM)))
```
Looking at marginal effects at 0 and 7 ACE scores to check for differences compared to model1 without interaction.
Marginal effects at 0 and 7 ACE score similar on ACE_CAT_SUM and SEXF. Though there was a statistically significant interaction effect, the size of the effect was not large enough to detect meaningful differences in MEs.

# plotting AMEs -----------------------------------------------------------
```{r, message=F}
library(sjPlot)
```
```{r}
plot_model(m1, type = "pred")

```





