---
title: "DEMassign1"
author: "Christopher Huong"
date: "`r Sys.Date()`"
output: html_document
---

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(car)
library(janitor)
library(ggplot2)
library(stargazer)
library(ggpubr)
library(ggrepel)
dat <- read_excel("Junkins Data.xlsx")
```
Load in relevant libraries and import excel data

```{r}
dat <- dat %>%
  mutate(south = if_else(region == 3, 1L, 0L),
         northeast = if_else(region == 1, 1L, 0L),
         midwest = if_else(region == 2, 1L, 0L),
         west = if_else(region == 4, 1L, 0L))

sum(dat$south, dat$midwest, dat$northeast, dat$west)
```
Create new variables 'south', 'northeast', 'midwest', and 'west'.
Assign south = 1 if region = 3, else south = 0. 1L means the integer 1, rather than 1.0000...
Assign northeast = 1 if region = 1, else northeast = 0
Assign midwest = 1 if region = 2, else midwest = 0
Assign west = 1 if region = 4, else west = 0

Then add the new columns to make sure they sum to 50

```{r}
dat <- dat %>%
  mutate(relcons2 = relcons^2,
         relcons3 = relcons^3,
         relconsln = log(relcons),
         relconsrec = relcons^-1)
```

Create new variables for transformed relcons

```{r}
plot(dat$relcons,dat$t_ageFM)

cor.test(dat$relcons,dat$t_ageFM)
```

Plot the scatter plot with x = relcons, and y = t_ageFM to visualize the relationship
Calculate the Pearson correlation between the variables
Notice a relatively strong negative relationship, with r=-0.57 confirming this

```{r}
mod1 <-lm(t_ageFM ~ relcons, dat)
summary(mod1)
```
Use OLS linear model to regress age of first marriage on religious conservatism
The model predicts for every 1 unit increase in relcons is associated with a -0.05 unit decrease in t_ageFM. ~32% of the variance of t_ageFM is explained by relcons. 


```{r}
mod2 <- lm(t_ageFM ~ relcons + relcons2, dat)
summary(mod2)
```
Add a quadratic relcons term to the regression to model a nonlinear relationship of y and x
No added variance is explained

```{r}
mod3 <-lm(t_ageFM~relcons + relcons2 + relcons3, dat)
summary(mod3)
```
Add a quadratic and cubic term to the regression model, no added variance is explained

```{r}
mod4 <-lm(t_ageFM~relconsln, dat)
summary(mod4)
```
Log-transforming the predictor, which should improve the model if the distribution of the predictor is skewed
No added variance is explained

```{R}
mod5 <-lm(t_ageFM~relconsrec, dat)
summary(mod5)
```
Taking the recipricol of the predictor
No added variance is explained


```{R}
BIC(mod1, mod2, mod3, mod4, mod5)
anova(mod1, mod2, mod3, mod4, mod5)
```
Directly comparing the models.

Model 1 has the lowest BIC, indicating that the relationship between t_ageFM and relcons is best explained by a linear model

```{R}
mod_med <-lm(t_ageFM~northeast + relcons, dat)

summary(mod_med)
```
Coefficient went from -0.04911 to -0.04375 and no added variance was explained, indicating that being a northeastern state does not mediate the relationship between t_ageFM and relcons

```{r}
mod_moder <- lm(t_ageFM~relcons*northeast, dat)
summary(mod_moder)
```
Modeling main effects of relcons and northeast, as well as their interaction
The northeast:relcons term is non-significant, indicating no interaction with relcons and northeast on t_ageFM

```{r}
mod_a <- lm(t_ageFM~northeast, dat)
mod_b <- lm (t_ageFM~northeast + relcons, dat)

stargazer(mod_a, mod_b,type="text", 
          column.labels = c("Model 1", "Model 2"), 
          intercept.bottom = FALSE, 
          single.row=FALSE,     
          notes.append = FALSE, 
          header=FALSE) 
```
Pretty table






















