---
title: "assign3"
output: html_document
date: "2023-02-12"
---




```{r}
data <- read.csv("junkinsdata.csv")


dat <- data[, c("ageFB", "t_ageFM", "nevermar", "cohabit", "abortion", "vryrel", "perBA")]

km2 <- kmeans(dat, 2, 1000)
```

Read in data, select variables we want to cluster by.
Perform k-means cluster analysis with clusters set to 2, to test the hypothesis that these variables reliably differ based on political orientation (democrat-republican), operationalized by state 2020 presidential election results. Iterations are set to 1000



```{r}
data$km2 <- km2$cluster
data$vote2020 <- c("R","R","R","D","D","D","D","D","R","D","D","R","R","D","R","R","R","D","D","D","D","D","R","R","R","R","R","R","D","D","D","D","D","R","R","D","D","D","R","R","R","R","R","R","D","D","D","D","R","R")
```
Add the cluster variable to the original data set 
Create new column with the 2020 presidential election results for each state 

```{r}
data <- data[order(data$km2),]
print(data[,c(1,64,63)])
table(data$vote2020, data$km2)
```
See how well k-means predicted state election vote by the selected variables, with k=2.
Performed OK but not great.

```{r}
pc <- prcomp(dat,
             center = T, scale. = T, retx = T)

dat1 <- cbind(data[,c(1,64,63)], data.frame(pc$x[,1]))
colnames(dat1) <- c("State", "Vote","Cluster", "PC1")
print(dat1[order(dat1$PC1),])
```
Conduct PCA on the same variables, and see if PCA predicts the state's cluster or vote.
Neither seem to be consistently associated with low/high PCA



The variables selected do not seem to reliably sort the states into republical/democrat.







